Coherent optical association of a single molecule

### 2020-11-04 paper.tex
Change tweezer powers

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{fig3.pdf}
  \caption{
    (a) Direct measurement of molecule lifetime in $15~\mathrm{mW}$ of trap depth.
    Molecule survival is detected by dissociating back to atoms using a second Raman transition.
    The lifetime is consistent with the $0.199(9)~\mathrm{ms}$
    measured from the Raman transition data.
    The oscillation in the survival is the result of the interference
    between the two Raman pulses with incomplete transfer.
    (b) Two-body atom lifetime of $5(1)~\mathrm{ms}$
    in $15~\mathrm{mW}$ of trap depth caused by off-resonance photoassociation.
    This is used to improve the fitting of the Raman transfer data.
    Inset: Atomic scattering rate scales as $P_{tweezer}^{2.58}$ on a log-log scale,
    this is consistent with a two photon scattering process.
    We have not measured a clear dependency of the loss rate on the tweezer detuning.
    \label{f-lifetime}}
\end{figure}

@@

We choose the tweezer frequency to be far detuned ($145 ~\mathrm{GHz}$) from the $v' = 0$ line, and guided by coupled channel calculations,
we locate the Raman resonance for the atom to molecule transition at
$770.57150(9) ~\mathrm{MHz}$ (Fig.~\ref{f-raman}c inset)
with $15 ~\mathrm{mW}$ tweezer power.
% (We can maybe add information about the prediction here?)
%% Prediction
% This excited state used in the Raman transition was measured in our previous experiment using photoassociation to be at $288560 GHz$ from our atomic state. The ground molecular state has not been observed previously in experimentally. Based on our measurement of FB resonance, interaction shift and the binding energy of the 4422 bound state. Theory prediction was at $770.1 MHz$.
% The background level of $31~\mathrm{\%}$ corresponds to the probability of preparing the two atoms in the relative motional ground state.
% When the atoms are transferred into the molecule state by the Raman transition, there is a decrease in the two body survival since the resulting molecule is dark to our imaging sequence. % directly detected by our imaging step.
The molecular state is dark to the imaging step,
so successful transfer of the atoms to the molecular state will be detected as loss.
We observed a Fourier limited linewidth which is evidence of a coherent transfer.
In order to verify the coherence of the transfer directly,
we fix the two photon detuning on resonance and scan the pulse time.
Fig.~\ref{f-raman}c shows the observed coherent Rabi oscillation between the atomic and molecular states.
Fitting the data with a decaying Rabi oscillation suggests that
$69~\mathrm{\%}$ of initial ground state atoms are transferred into the molecular state.
This transfer efficiency is mainly limited by the molecular lifetime
which can be measured directly by preparing the molecule with a $\pi $ pulse
and then using a second $\pi$ pulse to dissociate the molecule back to atoms
after a variable wait time (Fig.~\ref{f-lifetime}a).
The result shows a molecular lifetime of $0.199(9)~\mathrm{ms}$
consistent with the decay of Rabi oscillation.
We obtain the Raman Rabi frequency by fitting our measurements to a model
that includes a Raman Rabi frequency and
a finite lifetime for the molecular state (Fig.~\ref{f-raman}c and inset).
We account for the effect of atomic state loss by measuring the one and
two body lifetime of the atoms directly (Fig.~\ref{f-lifetime}b)
without turning on the second frequency.
The fit shows that we have a Raman Rabi frequency of $2\pi\times3.28(4) ~\mathrm{kHz}$.

@@

First we look at the change in resonance frequency.
As a function of the tweezer power,
we observe the expected linear dependency on the resonance frequency
caused by the differential light shift between the atomic and molecular state (Fig.~\ref{f-det}a).
When we vary the tweezer frequency around the $v'=0$ intermediate state,
we can further observe a $1/\Delta$ component
and a constant background in the experimentally explored region.
The background is caused by coupling to other excited states that are further away in energy.
The $1/\Delta$ component, however, is due to the coupling between the molecular state
and the $v'=0$ intermediate state.
From this measurement, we can extract a $\Omega_m$ of
$2\pi\times36.16(2) ~\mathrm{MHz}/\sqrt{\mathrm{mW}}$ or
$2\pi\times140.06(8) ~\mathrm{MHz}$ for the $15 ~\mathrm{mW}$ tweezer power used above.
This number is close to the value of
$2\pi\times27 ~\mathrm{MHz}/\sqrt{\mathrm{mW}}$ calculated from theory. \todo{ref/sm theory}

@@

In order to calculate the matrix element ratio,
we now need to extract the Rabi frequency between the atomic state and the intermediate state, $ \Omega_a $.
We do this by measuring the dependencies of the Raman Rabi frequency,
which depends on both $\Omega_m$ and $\Omega_a$.
The Raman Rabi frequency shows a non-linear dependency on the tweezer power
due to the change in the atomic wavefunction caused by
tighter confinement at higher power (Fig.~\ref{f-det}b).
As discussed before, for weakly interacting particles,
$\Omega_a$ scales as $ \omega_{\text{trap}}^{3/4}$ or $P^{0.375}$.
However, due to the strong interaction between the two atoms, this approximation breaks down.
Instead, coupled-channel calculations show that the scaling
is very well approximated by $P^{0.29}$ within the range of confinement in our experiment.
Combined with the standard intensity factor, the Raman Rabi frequency should scale as $P^{1.29}$,
which fits well to our experimental result.
Similar to the light shift, the Raman Rabi frequency's detuning dependence is determined by a constant background component and
a $v'=0$ component in the Raman Rabi frequency that scales as $1/\Delta$.
The $v'=0$ component of the Raman Rabi frequency is
$2\pi\times170(4)~\mathrm{Hz\cdot mW^{-1.29}}$,
or $2\pi \times 5.6(1)~\mathrm{kHz}$ at $15~\mathrm{mW}$ tweezer power.
Together with the $\Omega_m$ measured above, the matrix element, $ \Omega_a $, is
$2\pi \times 12.1(3)~\mathrm{MHz}$.
This gives a Rabi frequency, and therefore matrix element ratio of $11.6(3)$,
which is in fact better than the theory prediction of $23.7$.
Therefore, this should not cause the ratio of the Raman Rabi frequency to scattering rate
from the $v'=0$ state to be higher than expected.
Furthermore, we measure the natural linewidth of the $v'=0$ excited state to be no larger than $20~\mathrm{MHz}$ using photoassociation (PA) spectroscopy.
This suggests that the excited state linewidth should not cause
a stronger than expected scattering from $v'=0$ state either.

@@

\todo{change scattering to decoherence? since the fluctuation of light shift
  does not lead to scattering but only decherence.}
These results suggest that the decoherence or loss we observed during the Raman transition
comes from either a higher than expected background scattering rate
or a different intrinsic or technical source that we have not accounted for.
We have observed a significant decrease in the coherence time without the ASE filter,
suggesting the spectral purity of the laser is a significant source of scattering.
Other sources that can contribute to the decoherence includes
the stability of the tweezer power and the magnetic field.
Based on the Raman Rabi frequency to light shift ratio (Fig.~\ref{f-det}c),
the requirement on the tweezer power stability is $0.8~\mathrm{\%}$ at $15~\mathrm{mW}$.
We stabilize the tweezer power to $0.1~\mathrm{\%}$ so this should not be a major source
of decoherence.
Similarly, we measured a Zeeman shift of $42.2(2)~\mathrm{kHz/G}$
which does not cause significant decoherence from the measured magnetic field
fluctuation of $\sim1.5~\mathrm{mG}$.

In additional to calibrating the $\Omega_m$ and $\Omega_a$,
the scattering rate of the molecule depends on the tweezer power and detuning.
At $3~\mathrm{mW}$ tweezer power, we have observed a molecule lifetime as long as $1~\mathrm{ms}$.
However, since the technical noise that can lead to decoherence
is not fully characterized in our experiment,
we are unable to further identify the sources of the measured scattering rate
based on our measured detuning and power dependencies.

>>>>

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{fig3.pdf}
  \caption{
    (a) Direct measurement of molecule lifetime in $3.75~\mathrm{mW}$ of trap depth.
    Molecule survival is detected by dissociating back to atoms using a second Raman transition.
    The lifetime is consistent with the $0.199(9)~\mathrm{ms}$
    measured from the Raman transition data.
    The oscillation in the survival is the result of the interference
    between the two Raman pulses with incomplete transfer.
    (b) Two-body atom lifetime of $5(1)~\mathrm{ms}$
    in $3.75~\mathrm{mW}$ of trap depth caused by off-resonance photoassociation.
    This is used to improve the fitting of the Raman transfer data.
    Inset: Atomic scattering rate scales as $P_{tweezer}^{2.58}$ on a log-log scale,
    this is consistent with a two photon scattering process.
    We have not measured a clear dependency of the loss rate on the tweezer detuning.
    \label{f-lifetime}}
\end{figure}

@@

We choose the tweezer frequency to be far detuned ($145 ~\mathrm{GHz}$) from the $v' = 0$ line, and guided by coupled channel calculations,
we locate the Raman resonance for the atom to molecule transition at
$770.57150(9) ~\mathrm{MHz}$ (Fig.~\ref{f-raman}c inset)
with $3.75~\mathrm{mW}$ tweezer power.
% (We can maybe add information about the prediction here?)
%% Prediction
% This excited state used in the Raman transition was measured in our previous experiment using photoassociation to be at $288560 GHz$ from our atomic state. The ground molecular state has not been observed previously in experimentally. Based on our measurement of FB resonance, interaction shift and the binding energy of the 4422 bound state. Theory prediction was at $770.1 MHz$.
% The background level of $31~\mathrm{\%}$ corresponds to the probability of preparing the two atoms in the relative motional ground state.
% When the atoms are transferred into the molecule state by the Raman transition, there is a decrease in the two body survival since the resulting molecule is dark to our imaging sequence. % directly detected by our imaging step.
The molecular state is dark to the imaging step,
so successful transfer of the atoms to the molecular state will be detected as loss.
We observed a Fourier limited linewidth which is evidence of a coherent transfer.
In order to verify the coherence of the transfer directly,
we fix the two photon detuning on resonance and scan the pulse time.
Fig.~\ref{f-raman}c shows the observed coherent Rabi oscillation between the atomic and molecular states.
Fitting the data with a decaying Rabi oscillation suggests that
$69~\mathrm{\%}$ of initial ground state atoms are transferred into the molecular state.
This transfer efficiency is mainly limited by the molecular lifetime
which can be measured directly by preparing the molecule with a $\pi $ pulse
and then using a second $\pi$ pulse to dissociate the molecule back to atoms
after a variable wait time (Fig.~\ref{f-lifetime}a).
The result shows a molecular lifetime of $0.199(9)~\mathrm{ms}$
consistent with the decay of Rabi oscillation.
We obtain the Raman Rabi frequency by fitting our measurements to a model
that includes a Raman Rabi frequency and
a finite lifetime for the molecular state (Fig.~\ref{f-raman}c and inset).
We account for the effect of atomic state loss by measuring the one and
two body lifetime of the atoms directly (Fig.~\ref{f-lifetime}b)
without turning on the second frequency.
The fit shows that we have a Raman Rabi frequency of $2\pi\times3.28(4) ~\mathrm{kHz}$.

@@

First we look at the change in resonance frequency.
As a function of the tweezer power,
we observe the expected linear dependency on the resonance frequency
caused by the differential light shift between the atomic and molecular state (Fig.~\ref{f-det}a).
When we vary the tweezer frequency around the $v'=0$ intermediate state,
we can further observe a $1/\Delta$ component
and a constant background in the experimentally explored region.
The background is caused by coupling to other excited states that are further away in energy.
The $1/\Delta$ component, however, is due to the coupling between the molecular state
and the $v'=0$ intermediate state.
From this measurement, we can extract a $\Omega_m$ of
$2\pi\times72.32(4) ~\mathrm{MHz}/\sqrt{\mathrm{mW}}$ or
$2\pi\times140.06(8) ~\mathrm{MHz}$ for the $3.75~\mathrm{mW}$ tweezer power used above.
This number is close to the value of
$2\pi\times54 ~\mathrm{MHz}/\sqrt{\mathrm{mW}}$ calculated from theory. \todo{ref/sm theory}

@@

In order to calculate the matrix element ratio,
we now need to extract the Rabi frequency between the atomic state and the intermediate state, $ \Omega_a $.
We do this by measuring the dependencies of the Raman Rabi frequency,
which depends on both $\Omega_m$ and $\Omega_a$.
The Raman Rabi frequency shows a non-linear dependency on the tweezer power
due to the change in the atomic wavefunction caused by
tighter confinement at higher power (Fig.~\ref{f-det}b).
As discussed before, for weakly interacting particles,
$\Omega_a$ scales as $ \omega_{\text{trap}}^{3/4}$ or $P^{0.375}$.
However, due to the strong interaction between the two atoms, this approximation breaks down.
Instead, coupled-channel calculations show that the scaling
is very well approximated by $P^{0.29}$ within the range of confinement in our experiment.
Combined with the standard intensity factor, the Raman Rabi frequency should scale as $P^{1.29}$,
which fits well to our experimental result.
Similar to the light shift, the Raman Rabi frequency's detuning dependence is determined by a constant background component and
a $v'=0$ component in the Raman Rabi frequency that scales as $1/\Delta$.
The $v'=0$ component of the Raman Rabi frequency is
$2\pi\times1.02(2)~\mathrm{kHz\cdot mW^{-1.29}}$,
or $2\pi \times 5.6(1)~\mathrm{kHz}$ at $3.75~\mathrm{mW}$ tweezer power.
Together with the $\Omega_m$ measured above, the matrix element, $ \Omega_a $, is
$2\pi \times 12.1(3)~\mathrm{MHz}$.
This gives a Rabi frequency, and therefore matrix element ratio of $11.6(3)$,
which is in fact better than the theory prediction of $23.7$.
Therefore, this should not cause the ratio of the Raman Rabi frequency to scattering rate
from the $v'=0$ state to be higher than expected.
Furthermore, we measure the natural linewidth of the $v'=0$ excited state to be no larger than $20~\mathrm{MHz}$ using photoassociation (PA) spectroscopy.
This suggests that the excited state linewidth should not cause
a stronger than expected scattering from $v'=0$ state either.

@@

\todo{change scattering to decoherence? since the fluctuation of light shift
  does not lead to scattering but only decherence.}
These results suggest that the decoherence or loss we observed during the Raman transition
comes from either a higher than expected background scattering rate
or a different intrinsic or technical source that we have not accounted for.
We have observed a significant decrease in the coherence time without the ASE filter,
suggesting the spectral purity of the laser is a significant source of scattering.
Other sources that can contribute to the decoherence includes
the stability of the tweezer power and the magnetic field.
Based on the Raman Rabi frequency to light shift ratio (Fig.~\ref{f-det}c),
the requirement on the tweezer power stability is $0.8~\mathrm{\%}$ at $3.75~\mathrm{mW}$.
We stabilize the tweezer power to $0.1~\mathrm{\%}$ so this should not be a major source
of decoherence.
Similarly, we measured a Zeeman shift of $42.2(2)~\mathrm{kHz/G}$
which does not cause significant decoherence from the measured magnetic field
fluctuation of $\sim1.5~\mathrm{mG}$.

In additional to calibrating the $\Omega_m$ and $\Omega_a$,
the scattering rate of the molecule depends on the tweezer power and detuning.
At $0.75~\mathrm{mW}$ tweezer power, we have observed
a molecule lifetime as long as $1~\mathrm{ms}$.
However, since the technical noise that can lead to decoherence
is not fully characterized in our experiment,
we are unable to further identify the sources of the measured scattering rate
based on our measured detuning and power dependencies.
